Figure supplements

### 2022-12-02 text/Manuscript.Rmd
Figure_pdf

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva.(B-D) Expression analysis of the NOS gene (magenta) using in situ HCR. (D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). (B) Dorsal view of the larva at 3-day-old. (C and D) Anterior view of the larva at 2 (D) and 3-day-old (C). (E) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 3-day-old from the anterior side. (F and G) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (E) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. (H) Wiring diagram of the cPRC circuit.](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (B and C) Vertical swimming in wild type and mutant larvae at 2-day-old (B) and 3-day-old (C) stimulated with blue (488 nm) and UV (395 nm) light from above. The data are shown in 30 s bins. (D) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation. (E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0. (F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown. (E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (E and F) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae. (G-J) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig3.png){width="100%"}

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1). Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A). We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C). These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response. We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E). In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed. This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F). In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H). This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).

![**Figure 4. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) The intensity of DAF-FM in INNOS changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm). Hereafter, light purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean. (B and C) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (B: NOSΔ11/Δ11, B: NOSΔ23/Δ23) (light yellow) larvae. (D and G) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons. Correlation values were calculated relative to the cPRC (D) or IN-NOS (G). (E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae. (F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae. (H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae. (I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig4.png){width="100%"}

@@

>>>>

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A) SEM image of a 3-day-old Platynereis larva. (B and C) Five types of neurons (cPRC, INNOS, INRGWa, Ser-h1 and MC) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (C) are shown in purple. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D) Wiring diagram of the cPRC circuit. (E-G) Expression analysis of the NOS gene (magenta) using in situ HCR. (E) Dorsal view of the larva at 3-day-old. (F and G) Anterior view of the larva at 3 (F) and 2-day-old (G). (F) Insert: showing two NOS cells on each side. (G) Co-stained image using acetylated α-tubulin antibody (acTub: green). (H) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (NOSp::palmi-3xHA: magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia of cPRC and axonal scaffold (acTub: green). Views of larvae at 3-day-old from the anterior side.](../figures/Fig1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** 
(A) Vertical swimming in larvae treated with NOS inhibitors, L-NAME at 3-day-old stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins. (B) Top: The domain and Exon/Intron structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Gray shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild-type (WT) and mutant (NOSΔ11 and NOSΔ23) larvae at 2-day-old (C) and 3-day-old (D) stimulated with UV (395 nm) light from side, blue (488 nm) light from top and UV (395 nm) light from top. The data are shown in 30 s bins.

(E and F) Overlaid trajectories for WT (left, n=20) and NOS mutant (right, n=31) 3-day-old larvae. Grey shows the trajectories before UV stimulation and purple shows the trajectories after UV stimulation.

(E) The temporal changes in the vertical position of the WT and mutant larvae before and after UV stimulation are shown. The starting points of each larval trajectory are set to 0.

(F) The temporal changes in the distance traveled of the WT and mutant larvae before and after UV stimulation are shown.

(E and F) After UV stimulation is indicated by purple squares.](../figures/Fig2.png){width="100%"}

To test if NO is produced during UV/violet responses, we injected zygotes with the fluorescent NO-reporter DAF-FM and imaged two-day-old larvae during the local stimulation of the cPRC cilia by 405 nm light (Figure 4---figure supplement 1). Following light stimulation, we detected an increase in DAF-FM fluorescence in the region of the INNOS cells (Figure 4A). We next analysed the dynamics of the cPRC circuit following light stimulation. As we have shown previously [@Veraszto2018], a 20-sec illumination of the cPRC ciliary region lead to a decrease in cPRC calcium levels, likely due to the hyperpolarisation of the cPRCs. After 20-sec, calcium levels in cPRCs were raising again, reaching higher levels than at the start of the stimulus - a response that may involve depolarisation. This activation phase occurs after the end of the stimulation and may be due to a slower neuroendocrine feedback [@Veraszto2018]. To determine whether NO mediates such a feedback, we measured calcium signals in *NOS*-mutant larvae. While we detected the inhibition phase in cPRC in *NOS*-mutants, this was not followed by activation and calcium levels remained low (Figure 4B and C). These results revealed an essential requirement for NOS signalling in the second phase of the cPRC response.

![**Figure 3. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC.** (A) A NO-positive cell visualised by DAF-FM is shown (dashline). The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Intensity changes before and during 405 nm light stimulation in a cell visualised by DAF-FM. (C) The intensity of DAF-FM changes over time when cPRC (green) and other cells (NC) (gray) are stimulated with UV/violet (405 nm) light. Purple squares indicate the timing of 405 nm light stimulation, the transparent lines show the optical responses to individual intensities, each intensities are normalized (ΔF/F0) and the dark trace lines and the shaded bands show the mean value and the standard error range of the mean. (D and E) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in WT (dark yellow) and NOS mutant (D: NOSΔ11/Δ11, E: NOSΔ23/Δ23) (light yellow) larvae.](../figures/Fig3.png){width="100%"}

@@

![**Figure 4. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1.Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (B) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (C and D) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub).

(E-G) Changes over time in GCaMP6s intensities in cPRC when stimulated with UV in control (non-injection) and NIT-GC1 and NIT-GC2 knockdown by morpholino (MO1 and 2) larvae.

(H) The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. 
(I) A schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence.
(J-M) The intensity changes in cGMP fluorescence over time for the four conditions are shown. Gray traces show the optical responses to individual intensities (n \> 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars).](../figures/Fig4.png){width="100%"}

We next quantified the responses of the INNOS and INRGW interneurons following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method. We used the cell-specific neuropeptide markers anti-RYamide (INNOS) and anti-RGWa (INRGW) [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision (Figure 4---figure supplement 2). In both wild type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs (Figure 4D and E). In addition, In-NOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed. This indicates that cPRC hyperpolarisation is required for IN-NOS depolarisation (Figure 4F). In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants (Figure 4G and H). This suggests that there is no circuit output to the Ser-h1 cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects(Figure 4I).

![**Figure 5. NO produced by UV stimulation to cPRCs induces a depolarising response in cPRC via NIT-GC1.** (A) Schematic diagram of the immunostaining procedure after calcium imaging. (B and D) The position of the cPRC in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (C and E) Immunostaining image using RYamide (C) or RGWamide and 5-HT antibodies (red) (E) and DAPI (cyan) in the same region as (B) or (D). IN-NOS, IN-RGWa and Ser-h1 identified from the numbered nuclear positions are shown. (F) Correlation (Pearson's r) map of neuronal activity of the cPRC, IN-NOS, IN-RGWa and Ser-h1 neurons.

(E) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in WT (dark purple) and NOS mutant (light purple) larvae. (F) Changes over time in GCaMP6s intensities in IN-NOS when cPRC is stimulated with UV in NIT-GC2 morphants larvae. (H) Changes over time in GCaMP6s intensities in IN-RGWa when cPRC is stimulated with UV in WT (dark blue) and NOS mutant (light blue) larvae. (I) Changes over time in GCaMP6s intensities in Ser-h1 when cPRC is stimulated with UV in WT (dark red) and NOS mutant (light red) larvae.](../figures/Fig5.png){width="100%"}

@@

**NIT-GC Identification and Phylogenetic Analysis**
To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al. 2021). From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated. HMM models were run in HMMR3 with an e-value of 1e−15. The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15. We merged these two databases and ran CD-Hit (Fu et al. 2012) to eliminate redundant sequences (at a 99% threshold). To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig. S4, Supplementary Material online. To identify clusters, we used the convex-clustering option with 100 jack-knife replicates. The SREB receptors are extremely well conserved and form an easily recognizable cluster. To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building. We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al. 2009). To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al. 2020). The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online. The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.

**Single-cell analysis**
I used Achim et al 2015 for the single-cell data. In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells.
Since the raw data were read count data, I normalized them to TPM using Python.
After that we converted them to log10. From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell.
I chose the appropriate genes for now. For each cell, I identified them with marker genes.
After created the data in Python, plotted it using R dot plots.

RPKM? TPM?
RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples.
Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples.
In this case I wanted to compare between samples, so I used TPM.
In fact, TPM is becoming the mainstream method for many single-cell analyses.
The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.

### 2022-12-02 text/Supplements.Rmd
Figure_pdf

![**Figure 1---figure supplement 1.** (A) Phylogenetic tree of NOS by maximum likelihood (ML). Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig1_sup1.png){width="100%"}

![**Figure 1---figure supplement 2.** (A) Immunostaining with NOS antibodies.(B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva.](../figures/Fig1_sup2.png){width="100%"}

![**Figure 3---figure supplement 1.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum).](../figures/Fig3_sup1.png){width="100%"}

![**Figure 3---figure supplement 2.** (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. (B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig3_sup2.png){width="100%"}

![**Figure 3---figure supplement 3.** (A) Comparison of amino acid sequences of regions of possible NO binding in NIT-GC1 and NIT-GC2. (B) 3D stuructures of NIT-GC1 and NIT-GC2 (C)](../figures/Fig3_sup3.png){width="100%"}

![**Figure 3---figure supplement 4.** (A and B) Left: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Right shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence.](../figures/Fig3_sup4.png){width="100%"}

![**Figure 4---figure supplement 1.** (A) An INNOS visualised by DAF-FM is shown (a white arrowhead). A black arrowhead indicates the position of the adult eye. The white dashed line indicates the outer frame of the larva. All subsequent images are of 2-day-old larvae. (B) Dual-colour imaging of DAF-FM (green) and RGECO1a (magenta) allows simultaneous visualisation of the INNOS and cPRCs. (C) Enlarged view of (B). cPRCs and INNOS are indicated by arrows.](../figures/Fig4_sup1.png){width="100%"}

![**Figure 4---figure supplement 2.** (A) Schematic diagram of the immunostaining procedure after calcium imaging. (B) The position of the cPRC (purple dashed line) in calcium imaging with GCaMP6s. Cells around the cPRC are numbered according to the position of the nuclei. (C) Immunostaining image using RGWamide (RGWa) and serotonin (Ser) antibodies (red) and DAPI (light blue) in the same region as (B). INRGW (green dashed line) and Ser-h1 (black arrowheads) identified from the numbered nuclear positions are shown.](../figures/Fig4_sup2.png){width="100%"}

>>>>

![**Figure 1---figure supplement 1.** Phylogenetic tree of NOS by maximum likelihood. Tree robustness was tested with 1000 replicates of ultrafast bootstrap with the aLRT-SH-like and aBayes methods.](../figures/Fig1_sup1.png){width="100%"}

![**Figure 4---figure supplement 1.** Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclases (sGC) as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata.](../figures/Fig4_sup1.png){width="100%"}

![**Figure 4---figure supplement 2.** (A) Immunostaining with NOS antibodies. (B) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (A) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. (B) Co-localisation analysis using NIT-GC1 and NOS antibodies.](../figures/Fig4_sup2.png){width="100%"}
